---
title: "for loop practice"
author: "Sydney Rilum"
date: "2/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lets imagine that annual yield of a crop can be estimated follows:

yield=1.8∗fertilizer2−0.5∗fertilizer+0.3∗TP

and TP is mean precipitation in cm

Use a for loop to compute the total yield after 10 years:
- fertilizer is normally distributed with mean of 10 and standard deviation of 0.5
- TP is 20cm

```{r}
# make a random sample of fertilizer application, 10 values with mean of 5 and standard deviation of 0.5
fertilizer = rnorm(n = 10, mean = 5, sd = 0.5)

# create a data frame to store results
results = data.frame(fertilizer = fertilizer, crop_yield = NA)

# TP is mean precipitation in cm
TP = 20

# use a for loop to compute the total yield after 10 years
for (i in 1:10) {
  results$crop_yield[i] = 1.8 * results$fertilizer[i]^2 - 0.5 * results$fertilizer[i] + 0.3 * TP
}

```

Now create a function to that takes as input a single value for annual fertilizer and annual total precipitation and returns yield
```{r}
# crop yield function 
crop_yield = function(fertilizer, TP){
  yield = 1.8 * fertilizer^2 - 0.5 * fertilizer + 0.3 * TP
  return(yield)
}  

# apply function
crop_yield(fertilizer = c(1, 2, 3), TP = c(10, 20, 30))
```

Now modify your function so that it returns total yield if the user inputs a vector of fertilizer and a vector of precipitation
```{r}
# crop yield function 
crop_yield = function(fertilizer, TP){
  yield = 1.8 * fertilizer^2 - 0.5 * fertilizer + 0.3 * TP
  total_yield = sum(yield)
  return(total_yield)
}  

# use total yield
crop_yield(fertilizer = c(1, 2, 3), TP = c(10, 20, 30))
```

Check that function returns the same result as your original for loop Add some error checking and test
```{r}
crop_yield = function(fertilizer, TP){
  # error checking
  fetilizer = ifelse(fertilizer < 0, return("fertilizer cannot be negative"), fertilizer)
  # function
  yield = 1.8 * fertilizer^2 - 0.5 * fertilizer + 0.3 * TP
  return(yield)
}  

# test error checking 
crop_yield(fertilizer = -2, TP = 20)
```


Note: you can put for loops inside of functions
Note: for loops can also call functions






